\begin{thebibliography}{99}
\bibitem{bKukhtarev_phr} N.V. Kukhtarev, V.B. Markov, 
S.G. Odulov, M.S. Soskin and
V.L. Vinetskii. {\it Holographic storage in electrooptic crystals. I. Steady
state.} Ferroelectrics, v 22, pp. 949-960, 1979  
\bibitem{bGunter_phr} George C. Valley, Juan F. Lam. {\it Theory of photorefractive
effects in electro-optic crystals.} Topic in Applied Physics v.61
Photorefractive Materials and their applications I. Findamental phenomena, p.
75  Springer-Verlag Berlin Heidelberg, 1988, pp. 75-98 
\bibitem{bValley_phr} George C. Valley, Martin B. Klein {\it
Optimal properties of photorefractive materials for optical data processing},
Optical Engineering, v. 22(6), pp. 704-711, 1983
\bibitem{bPetrovNonStat_phr} Sergei I. Stepanov, Mikhail P. Petrov {\it Nonstationary 
holographic recording for efficient amplification and phase conjugation.} 
Topic in Applied Physics v.61
Photorefractive Materials and their applications I. Findamental phenomena, p.
75  Springer-Verlag Berlin Heidelberg, 1988, pp. 263-289 
\bibitem{bND_phr} Peltier M., Micheron F.  J. Appl.
Physics, 1977, v. 48, p. 3683
\bibitem{bNA_phr} Feinberg J., Heiman D., Tanguary A. R., Hellwarth R. W. J. Appl.
Physics, 1980, v. 51, p. 1297
\bibitem{bEpsilon_phr} Gunter P.  Physics Reports, 1983, v. 9, p. 200
\bibitem{bPetrov} М.П. Петров, С.И. Степанов, А.В. Хоменко. {\it
Фоторефрактивные кристаллы в когерентной оптике.} С. Петербург, Наука, 1992.
\bibitem{bIntEq1} Петрунькин В.Ю., Водоватов И.А. Известия ВУЗов. 
Радиофизика, 1983, т. 26, с. 1570
\bibitem{bIntEq2} Водоватов И.А., Плисс Н.С., Попова Л.П., Пучкова А.И. 
 Научное приборостроение, 1991, н. 2, с.165
\bibitem{bIntEqPhr1} Водоватов И.А., Плисс Н.С., Попова Л.П., Пучкова А.И.
Радиотехника и электроника, 1994, т. 37, с. 1939
\bibitem{bWeinstein} П. А. Вайнштейн {\it Электромагнитные волны.} М: 
Радио и связь, 1988. 
\end{thebibliography}
