\begin{thebibliography}{99} 
\bibitem{OurDPCM} Письма в ЖТФ
\bibitem{Gunter} George C. Valley, Juan F. Lam. {\it Theory of photorefractive
effects in electro-optic crystals.} Topic in Applied Physics v.61
Photorefractive Materials and their applications I. Findamental phenomena, p.
75  Springer-Verlag Berlin Heidelberg, 1988 
\bibitem{Psaltis} D.Psaltis, D.Brady, K.Wagner 
{\it Adaptive optical networks using photorefractive crystals} Applied
Optics, Vol. {\bf 27(9)}, pp.1752-1759, 1988.
\bibitem{ART} D.C.Wunsch, D.J.Morris et al. {\it Photorefractive adaptive resonans neural
network} Appl.Optics, {\bf 32} No~8, 1993.
\bibitem{Time-dependent} M.Horowitz, D.Kligler, B.Fischer 
{\it Time-dependent behavior of photorefractive two- and four-wave 
mixing.} J. of the Opt. Society of America B, {\bf 8(10)}, 1991
\bibitem{competition} Q.Byron He, Pochi Yeh, Claire Gu, R.R.
Neurgaonkar {\it Multigrating competition effects in
photorefractive mutually pumped phase conjugation} J. of the
Optical Society of America B, {\bf 9(1)}, 1992.
\bibitem{Dynamics}  М.П.Петров, Х.Д.Колфилд, Е.В.Мокрушина {\it Некогерентное двойное обращающее зеркало на
$Bi_{12}TiO_{20}$} Квантовая электроника, {\bf 19}, No~3, 1992.
\bibitem{Weinstein} П. А. Вайнштейн {\it Электромагнитные волны.} М: 
Радио и связь, 1988.  
\bibitem{material} М. П. Петров, С. И. Степанов, А. В. Хоменко
{\it Фоторефрактивные кристаллы в когерентной оптике} СПб, Наука, 1992
\bibitem{Zozulya} А. А. Зозуля {\it Влияние фенинга на характеристики
самонакачивающихся систем четырехволнового смешения в фоторефрактивных
средах,} Квантовая электроника, {\bf 19}, N 8, 1992   

\end{thebibliography} 