{\small
\
\begin{thebibliography}{99}

\bibitem{Horowitz1991}M.Horowitz, D.Kligler, B.Fisher,{
"Time-dependent behavior of photorefractive two- and four-wave
mixing"}, J.of Opt.Soc.B, {\bf 8(10)}, 2204-2217, 1991.
\bibitem{Bel'dyugina1992} N.I.Bel'dyugina, A.V.Mamayev,
V.V.Shkunov, "Self-starting generation dynamics in the semilinear
phase conjugate mirror with a large number of the transverse
modes.", {\it Kvantovaya Elektronika}, {\bf 19},691-697, 1992.
\bibitem{Dynamics} M.P.Petrov, H.D.Colfield, E.V.Mokrushina
{"Noncoherent double conjugate mirror using $Bi_{12}TiO_{20}$"}
Proc. of SPIE, v.1731, pp. 18-29(1991).
\bibitem{Psaltis} D.Psaltis, D.Brady, K.Wagner {"Adaptive
optical networks using photorefractive crystals"} Applied Optics,
Vol. {\bf 27(9)}, pp.1752-1759, (1988).
 \bibitem{ART} D.C.Wunsch, D.J.Morris et al. {"Photorefractive adaptive
  resonans neural 
network,"} Appl.Optics, {\bf 32} No~8, (1993).
%\bibitem{oscillators} B.Fischer, S.Sternklar, S.Weiss {"Photorefractive
%oscillators"}, IEEE Journal of Quantum Electronics, {\bf 25} No 3,
%(1989).
%\bibitem{attard} Alfred E.Attard, "Theory and origins of the
%photorefractive effects in $Bi_{12}SiO_20$, J. of Appl. Phys. {\bf
%69(1)}, pp.44-55 (1991).
%\bibitem{Sternklar1986} S.Sternklar, S.Weiss, M.Segev, B.Fischer,
%"Beam coupling and locking of lasers using photorefractive
%four-wave mixing", {\it Opt.Lett}, {\bf 11}, 528-530, 1986.
%\bibitem{Zhanuzakov1990} M.G.Zhanuzakov,A.A.Zozulya,V.T.Tikhnochuk,
%\it{"Stability of the steady nonlinear states of the light waves
%in the system of two coupled double conjugate mirrors with
%photorefractive crystals"}, Kvantovaya Elektronika, {\bf 17},
%1990.
%\bibitem{competition} Q.Byron He, Pochi Yeh, Claire Gu, R.R.
%Neurgaonkar {\it Multigrating competition effects in
%photorefractive mutually pumped phase conjugation} J. of the
%Optical Society of America B, {\bf 9(1)}, 1992.
\bibitem{bFirstSemi}  Cronin-Golomb M., Fisher B., Nilsen J. et al.
{"Laser with dynamic holographic intracavity distortion correction
capability."}  Applied Physics Letters. {\bf
  41(3)}. pp. 219-220. (1982)
\bibitem{Zozulya} A.A.Zozulya {"Influence of fanning on the features
of the our-wave mixing self-pumping systems in the photorefracieve
media,"} Kvantovaya Elektronika, {\bf 19}, No~8, (1992) (in
Russian).
\bibitem{ourDPCM} F.N. Nikiforov, I.V. Murashko,
I.A. Vodovatov, V.Yu. Petrun'kin, E.V. Mokrushina, "The Dynamics
of Double Reversal of Wave Fronts in Photorefractive Crystals,"
Technical Physics Letters, {\bf 26}, pp. 81-84 (2000).
\bibitem{ourSemi} Murashko, Ivan V., Nikiforov, P. N., Petrun'kin, Vsevolod Y., Vodovatov, Igor A. 
Light self-diffraction in semilinear generator geometry 
Proc. of SPIE, {\bf 4680}, pp. 224-232 (2002).
\end{thebibliography}
} %/small
